---
title: "blblm-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{blblm-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(blblm)
```

# Bag of Little Bootstraps Algorithm
The Bag of Little Bootstraps Algorithm works by splitting the data into subsamples a given number of times. Then, for each subsample, a bootstrap sample is taken. For each bootstrap sample, summary statistics of the model (coefficients and sigma for linear regression, coefficients for logistic regression) are computed. Then, the average is taken for each summary statistic to estimate the summary statistics of the data.

This R package implements the Bag of Little Bootstraps Algorithm in linear and logistic regression. The functions available to be used are detailed in the sections below.

# Functions for Linear Regression
Users have the option to use the function blblm or future_blblm to implement the Bag of Little Bootstraps algorithm in linear regression. The difference between blblm and future_blblm is that future_blblm uses parallelization via the R package "furrr."

## blblm
This function fits a linear model on a given data set using the bag of little bootstraps algorithm.

Input:

- formula: A symbolic description of the linear regression model to be fitted.

- data: A data set, such as a data frame or list, containing the variables in the model.

- m: Number indicating how many sub-samples the data will be split into. Default is 10.

- B: Number of bootstrap samples. Default is 5000.

Return:

Coefficients of each fit.

```{r}
blblm <- function(formula, data, m = 10, B = 5000) {
  data_list <- split_data(data, m)
  estimates <- map(
    data_list,
    ~ lm_each_subsample(formula = formula, data = ., n = nrow(data), B = B)
  )
  res <- list(estimates = estimates, formula = formula)
  class(res) <- "blblm"
  invisible(res)
}
```

## future_blblm
This function fits a linear model on a given data set using the bag of little bootstraps algorithm and parallelization for efficiency.

Input:

- formula: A symbolic description of the linear regression model to be fitted.

- data: A data set, such as a data frame or list, containing the variables in the model.

- m: Number indicating how many sub-samples the data will be split into. Default is 10.

- B: Number of bootstrap samples. Default is 5000.

- w: Number of workers. Default is 4.

Return:

Coefficients of each fit.

```{r}
future_blblm <- function(formula, data, m = 10, B = 5000, w = 4) {
  library(furrr)
  plan(multiprocess, workers = w)
  options(future.rng.onMisuse = "ignore")

  data_list <- split_data(data, m)

  estimates <- future_map(
    data_list,
    ~ lm_each_subsample(formula = formula, data = ., n = nrow(data), B = B)
  )
  res <- list(estimates = estimates, formula = formula)
  class(res) <- "blblm"
  invisible(res)
}
```


# Functions for Logistic Regression
Users have the option to use the function blblogreg or future_blblogreg to implement the Bag of Little Bootstraps algorithm in logistic regression. The difference between blblog and future_blblogreg is that future_blblogreg uses parallelization via the R package "furrr."

## blblogreg
This function fits a logistic regression model on a given data set using the bag of little bootstraps algorithm.

Inputs:

- formula: A symbolic description of the logistic regression model to be fitted.

- data: A data set, such as a data frame or list, containing the variables in the model.

- m: Number indicating how many sub-samples the data will be split into. Default is 10.

- B: Number of bootstrap samples. Default is 5000.

Return:

Coefficients of each fit.

```{r}
blblogreg <- function(formula, data, m = 10, B = 5000) {
  data_list <- split_data(data, m)
  estimates <- map(
    data_list,
    ~ glm_each_subsample(formula = formula, data = ., n = nrow(data), B = B)
  )
  res <- list(estimates = estimates, formula = formula)
  class(res) <- "blblm"
  invisible(res)
}
```

## future_blblogreg
This function fits a logistic regression model on a given data set using the bag of little bootstraps algorithm and parallelization for efficiency.

Inputs:

- formula: A symbolic description of the logistic regression model to be fitted.

- data: A data set, such as a data frame or list, containing the variables in the model.

- m: Number indicating how many sub-samples the data will be split into. Default is 10.

- B: Number of bootstrap samples. Default is 5000.

- w: Number of workers. Default is 4.

Return:

Coefficients of each fit.

```{r}
future_blblogreg <- function(formula, data, m = 10, B = 5000, w = 4) {
  library(furrr)
  suppressWarnings(plan(multiprocess, workers = w))
  options(future.rng.onMisuse = "ignore")

  data_list <- split_data(data, m)

  estimates <- future_map(
    data_list,
    ~ glm_each_subsample(formula = formula, data = ., n = nrow(data), B = B)
  )
  res <- list(estimates = estimates, formula = formula)
  class(res) <- "blblm"
  invisible(res)
}
```

# Supporting Functions
The following functions can be used on fitted blblm or blblogreg models unless otherwise specified to retrieve specific values/estimates.

## coef

The following functions estimates the coefficients for a fitted blblm or blblogreg model.

Inputs:

- object: Fitted blblm or blblogreg model.

- ...: Additional arguments.

Return: 
Estimated coefficients of the given fitted blblm or blblogreg model.

```{r}
coef.blblm <- function(object, ...) {
  est <- object$estimates
  map_mean(est, ~ map_cbind(., "coef") %>% rowMeans())
}

coef.blblogreg <- function(object, ...) {
  est <- object$estimates
  map_mean(est, ~ map_cbind(., "coef") %>% rowMeans())
}
```

## sigma
This function estimates sigma and, optionally, the confidence interval for sigma for a fitted blblm model. This function can only be used on blblm objects.

Inputs:

- object: Fitted blblm model.

- confidence: Single logical indicating whether the output should contain the confidence interval.

- level: Confidence level.

- ...: Additional arguments.

Return: 

Estimated value for sigma of the given fitted blblm model and, optionally, the confidence interval for sigma at the specified confidence level.

```{r}
sigma.blblm <- function(object, confidence = FALSE, level = 0.95, ...) {
  est <- object$estimates
  sigma <- mean(map_dbl(est, ~ mean(map_dbl(., "sigma"))))
  if (confidence) {
    alpha <- 1 - 0.95
    limits <- est %>%
      map_mean(~ quantile(map_dbl(., "sigma"), c(alpha / 2, 1 - alpha / 2))) %>%
      set_names(NULL)
    return(c(sigma = sigma, lwr = limits[1], upr = limits[2]))
  } else {
    return(sigma)
  }
}
```

## confint
This function computes confidence interval(s) for blblm or blblogreg model estimate(s) or given parameter(s) at a specified confidence level.

Inputs:

- object: Fitted blblm or blblogreg model.

- parm: Parameters to calculate confidence intervals for.

- level: Confidence level.

- ...: Additional arguments.

Return:

Confidence interval(s) for blblm or blblogreg model estimate(s) or given parameter(s) at a specified confidence level.

```{r}
confint.blblm <- function(object, parm = NULL, level = 0.95, ...) {
  if (is.null(parm)) {
    parm <- attr(terms(object$formula), "term.labels")
  }
  alpha <- 1 - level
  est <- object$estimates
  out <- map_rbind(parm, function(p) {
    map_mean(est, ~ map_dbl(., list("coef", p)) %>% quantile(c(alpha / 2, 1 - alpha / 2)))
  })
  if (is.vector(out)) {
    out <- as.matrix(t(out))
  }
  dimnames(out)[[1]] <- parm
  out
}

confint.blblogreg <- function(object, parm = NULL, level = 0.95, ...) {
  if (is.null(parm)) {
    parm <- attr(terms(object$formula), "term.labels")
  }
  alpha <- 1 - level
  est <- object$estimates
  out <- map_rbind(parm, function(p) {
    map_mean(est, ~ map_dbl(., list("coef", p)) %>% quantile(c(alpha / 2, 1 - alpha / 2)))
  })
  if (is.vector(out)) {
    out <- as.matrix(t(out))
  }
  dimnames(out)[[1]] <- parm
  out
}
```

## predict
This function predicts new values given new data using based on a fitted blblm or blblogreg model. 

Inputs:

- object: Fitted blblm or blblogreg model.

- parm: Parameters to be estimated with confidence intervals.

- level: Confidence level.

- ...: Additional arguments.

Return:

New, predicted values based on given new data and a given blblm or blblogreg model. 

```{r}
predict.blblm <- function(object, new_data, confidence = FALSE, level = 0.95, ...) {
  est <- object$estimates
  X <- model.matrix(reformulate(attr(terms(object$formula), "term.labels")), new_data)
  if (confidence) {
    map_mean(est, ~ map_cbind(., ~ X %*% .$coef) %>%
      apply(1, mean_lwr_upr, level = level) %>%
      t())
  } else {
    map_mean(est, ~ map_cbind(., ~ X %*% .$coef) %>% rowMeans())
  }
}

predict.blblogreg <- function(object, new_data, confidence = FALSE, level = 0.95, ...) {
  est <- object$estimates
  X <- model.matrix(reformulate(attr(terms(object$formula), "term.labels")), new_data)
  if (confidence) {
    map_mean(est, ~ map_cbind(., ~ X %*% .$coef) %>%
      apply(1, mean_lwr_upr, level = level) %>%
      t())
  } else {
    map_mean(est, ~ map_cbind(., ~ X %*% .$coef) %>% rowMeans())
  }
}
```

## Examples

Examples using all of the above functions can be found in the README.
